---
title: email
output: blastula::blastula_email
---

<style type = "text/css">

h1, h2, h3 {
  font-family: Arvo, "Helvetica Neue", Helvetica, Arial, sans-serif;
  font-weight: normal
}

p {
  font-family: "Source Sans Pro", "Helvetica Neue", Helvetica, Arial, sans-serif
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fontfamily = "Sans Source Pro",
  results="asis",
  fig.align='center',
  fig.cap = "" 
)
```



```{r, class.output = "banner", echo=FALSE, include =FALSE}
es <- azure_endpoint_url()
# storage endpoint
se <- AzureStor$storage_endpoint(es, sas = Sys.getenv("DSCI_AZ_SAS_DEV") )

# storage container
sc_projects <-  AzureStor$storage_container(se, "projects")

# Load thresholds data.frame ----------------------------------------------
tf <- tempfile(fileext = ".png")

invisible(AzureStor$download_blob(
  container = sc_projects, 
  src = "ds-contingency-pak-floods/centre_banner.png",
  dest = tf
)
)
```

```{r addBanner}
blastula$add_image(
  file = tf,
  alt = "Centre for Humanitarian Data banner",
  align = "left",
  width = "25%"
)
```


# Pakistan Rainfall Monitoring 2024

## Lower Indus Basin - `r email_txt$status`

### `r format(Sys.Date(),'%e %B %Y')`

This analysis is run daily to calculate the previous 3 day cumulative rainfall over the Lower Indus Basin in Pakistan. The monitoring occurs at 3 pm each day when the global IMERG data set for the previous day is published.


```{r}
btools$add_ggplot_custom(
  plot_object = p,
  alt = "Alert Map",
  height = 3.75,
  width = 5.5,
  html_width = 650
)
```


    
**Data source:**  [IMERG V7](https://gpm.nasa.gov/resources/documents/imerg-v07-atbd)

Data accessed `r trimws(format(as.Date(Sys.Date()), "%e %B %Y"))`.

Full documentation and source code can be found in the [GitHub Repository](https://github.com/OCHA-DAP/ds-pak-flooding-contingency)

Additional exploration and methodology notes can also be found [here](https://rpubs.com/zackarno/1199575)

----

## Contact

zachary.arno@un.org

----

```{r downloadFooter, echo=FALSE, include=F}

tf <- tempfile(fileext = ".png")

capture.output(
  invisible(
  AzureStor$download_blob(
  container = sc_projects, 
  src = "ds-contingency-pak-floods/ocha_logo_wide.png",
  dest = tf
)
)
)
```

```{r}
blastula$add_image(
  file = tf,
  alt = "Centre for Humanitarian Data banner",
  align = "center",
  width = "25%"
)
```

<p>
<center>

<b> OCHA Centre For Humanitarian Data </b>

Fluwelen Burgwal 58 | 2511 CJ The Hague | The Netherlands

</center>
</p>
